Smart Spreading Factor Assignment for LoRaWANs

### 2019-01-18 paper.tex
spell check

\par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between receive sensitivity and air time of transmission \cite{7803607}. Simultaneous same SF transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other. Thus SF assignment is crucial for overall network performance. In this work, a LoRa discrete event simulator is developed to study the performance of LoRa SF assignment strategies. Also a machine learning based SF assignment approach is proposed. Support Vector Machine (SVM) and Decision Tree Classifier (DTC) techniques are employed and the introduced schemes are called as smart SF schemes. The performance of the smart schemes  is compared with the performance of the lowest SF assignment scheme. It is shown that the proposed smart SF assignment schemes give promising results.

\par This paper is organized as follows: Section \ref{LoRa} and Section \ref{LoRaWAN} provides some background information about LoRa and LoRaWAN. Section \ref{Related Works} summarize  the other related works. Section \ref{Proposed Technique} describes the smart SF assignment technique proposed. Simulation environments and results are shown in Section \ref{Simulation Environment} and \ref{Simulation Results}. Finally, Section \ref{Conclusion} concludes the paper by givin future directions.

@@

\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single GW is illustrated. Different color rings represent achievable range of different SFs from the GW and different color circles represent selected SF of the nodes. The end devices close to the GW will fall into lowest SF (SF7) area section. The end devices close to the GW will probably select the lowest SF mot the time.Hence the number of collisions will increase as the number of end devices close to the GW increases.

@@

\par In \cite{8267219}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions. The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal significantly is greater than the reference signal. Their experimental results show that this power difference is around 16 dB such a this power difference can be seen when an interfering signal is close to a receiver or the sum of interfering signals' energy can create this power difference.

@@

\par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes' SFs requires different rules. For this reason, machine learning based SF assignment approach is proposed to decrease the colisioins for the same SF transmissions. This technique starts by learning the transmission behavior of the nodes in a network. An NS can keep track of successful uplink transmissions and their SFs. This NS can also keep track of some of the collided transmissions if header part of the packet is not interfered at the GW. However, NS cannot keep track of transmissions with lower receive power than sensitivity of the GW. Using this information, NS can train a classifier to predict future transmission result for a specific node and a specific SF. Using this prediction model NS can assign SFs to nodes considering the collision probability.

@@

\par It is possible to generate mass amount of LoRaWAN transmission records for different topologies using our simulator. Definetely, data set size is directly proportional to simulation duration. Thus, increasing the simulation duration, improves the prediction accuracy. In real world deployments, NS can keep track of all transmissions and it can create a classifier daily basis, then GWs can request from nodes to use suggested SFs.

@@

\par In Figure \ref{fig:sf_pdr}, PDR plot of various SF assignment is shown. Randomly generated network topology radius is set to 3000 meters and number of GWs is set to 1. Increasing SFs, increases air time. It increases the number of collisions thus decreases the PDR. High SF schemes gives poor PDR results as the number of nodes increases. Since network topology radius is quite small, all SFs can reach to GW.

@@

\par PDR plot of the lowest SF scheme and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}. Randomly generated network radius is set to 5000 meters and number of GWs is set to 3. Prediction model needs nodes' locations and three GWs are enough to locate nodes positions by triangulation. In Table \ref{table:prediction_pdr}, PDR values for various network radii are presented. Both smart SVM and smart DTC schemes give better PDR than lowest SF schemes when number of nodes increases. Increasing number of nodes, increases number of interference. Smart schemes  improve network performance when LoRa interference is high. Moreover, smart schemes give better results when nodes are deployed closer to the GW, since nodes have margin to increase their SFs when they are deployed closer to the GW. If a node is far away from GW, then smart schemes cannot increase the SF to avoid interference since the asigned SF is already high.

\par In Table \ref{table:prediction_accuracy}, prediction accuracy of smart SVM and smart DTC schemes are presented for various network radii and number of nodes. Prediction accuracy is not directly proportional to network PDR. Correct prediction of a interfered transmission may not increase the PDR but increases the accuracy. Mart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high.

@@

\par For space constraints, we omit results like network throughput, total transmit energy consumption. We invite readers to experiment the simulation tool with the parameters they like. \cite{simlorasf}.

\section{Conclusion} \label{Conclusion}
\par In this paper, we present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes. Moreover, we show how SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme. We present simulation results for the lowest SF assignment scheme and the proposed schemes. Simulation results show that, the the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to GWs.

>>>>

\par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between receive sensitivity and air time of transmission \cite{7803607}. Simultaneous same SF transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other. Thus, SF assignment is crucial for overall network performance. In this work, a LoRa discrete event simulator is developed to study the performance of LoRa SF assignment strategies. Also, a machine learning based SF assignment approach is proposed. Support Vector Machine (SVM) and Decision Tree Classifier (DTC) techniques are employed and the introduced schemes are called as smart SF schemes. The performance of the smart schemes is compared with the performance of the lowest SF assignment scheme. It is shown that the proposed smart SF assignment schemes give promising results.

\par This paper is organized as follows: Section \ref{LoRa} and Section \ref{LoRaWAN} provides some background information about LoRa and LoRaWAN. Section \ref{Related Works} summarize  the other related works. Section \ref{Proposed Technique} describes the smart SF assignment technique proposed. Simulation environments and results are shown in Section \ref{Simulation Environment} and \ref{Simulation Results}. Finally, Section \ref{Conclusion} concludes the paper by giving future directions.

@@

\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single GW is illustrated. Different color rings represent achievable range of different SFs from the GW and different color circles represent selected SF of the nodes. The end devices close to the GW will fall into lowest SF (SF7) area section. The end devices close to the GW will probably select the lowest SF mot the time. Hence the number of collisions will increase as the number of end devices close to the GW increases.

@@

\par In \cite{8267219}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions. The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal significantly is greater than the reference signal. Their experimental results show that this power difference is around 16 dB such a power difference can be seen when an interfering signal is close to a receiver or the sum of interfering signals' energy can create this power difference.

@@

\par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes' SFs requires different rules. For this reason, machine learning based SF assignment approach is proposed to decrease the collisions for the same SF transmissions. This technique starts by learning the transmission behavior of the nodes in a network. An NS can keep track of successful uplink transmissions and their SFs. This NS can also keep track of some of the collided transmissions if header part of the packet is not interfered at the GW. However, NS cannot keep track of transmissions with lower receive power than sensitivity of the GW. Using this information, NS can train a classifier to predict future transmission result for a specific node and a specific SF. Using this prediction model NS can assign SFs to nodes considering the collision probability.

@@

\par It is possible to generate mass amount of LoRaWAN transmission records for different topologies using our simulator. Data set size is directly proportional to simulation duration. Thus, increasing the simulation duration, improves the prediction accuracy. In real world deployments, NS can keep track of all transmissions and it can create a classifier daily basis, then GWs can request from nodes to use suggested SFs.

@@

\par In Figure \ref{fig:sf_pdr}, PDR plot of various SF assignment is shown. Randomly generated network topology radius is set to 3000 meters and number of GWs is set to 1. Increasing SFs increases air time. It increases the number of collisions thus decreases the PDR. High SF schemes gives poor PDR results as the number of nodes increases. Since network topology radius is quite small, all SFs can reach to GW.

@@

\par PDR plot of the lowest SF scheme and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}. Randomly generated network radius is set to 5000 meters and number of GWs is set to 3. Prediction model needs nodes' locations and three GWs are enough to locate nodes positions by triangulation. In Table \ref{table:prediction_pdr}, PDR values for various network radii are presented. Both smart SVM and smart DTC schemes give better PDR than lowest SF schemes when number of nodes increases. Increasing number of nodes, increases number of interferences. Smart schemes improve network performance when LoRa interference is high. Moreover, smart schemes give better results when nodes are deployed closer to the GW, since nodes have margin to increase their SFs when they are deployed closer to the GW. If a node is far away from GW, then smart schemes cannot increase the SF to avoid interference since the assigned SF is already high.

\par In Table \ref{table:prediction_accuracy}, prediction accuracy of smart SVM and smart DTC schemes are presented for various network radii and number of nodes. Prediction accuracy is not directly proportional to network PDR. Correct prediction of an interfered transmission may not increase the PDR but increases the accuracy. Mart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high.

@@

\par For space constraints, we omit results like network throughput and total transmit energy consumption. We invite readers to experiment the simulation tool with the parameters they like. \cite{simlorasf}.

\section{Conclusion} \label{Conclusion}
\par In this paper, we present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes. Moreover, we show how SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme. We present simulation results for the lowest SF assignment scheme and the proposed schemes. Simulation results show that, the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to GWs.
